### 2023-10-30 draft_jss/draft_jss.Rmd
update v0.0.1

@@

RC exhibits several advantages over its counterparts. It excels in tasks
that require the processing of temporal data, such as time series
forecasting and speech recognition, thanks to its inherent memory
capabilities. RC offers computational efficiency, reduced susceptibility
to overfitting, and a higher degree of interpretability compared to deep
learning models. Moreover, RC is able to capture intricate temporal
patterns compared to statistical methods such as penalized regression.

@@

TODO: - Ghosh S, Senapati A, Mishra A, Chattopadhyay J, Dana SK, Hens C,
et al. Reservoir computing on epidemic spreading: A case study on
COVID-19 cases. Phys Rev E. 16 juill 2021;104(1):014308. - Existing
review?

@@

RC use is stille challenging for several reasons. First, its stochastic
nature due to the reservoir random part makes the results of a given RC
unpredictable even if the hyperparameters are well chosen. Second,
eventhough, RC relies on several hyperparameter to determine the
connections between neurons. There exist few guidance about the way to
set the hyperparameters and most often the user must rely on a wrapper
approach were combinations of hyperparameters performance are evaluated
on a train set before chosing the ideal combination one. Third, most
published paper are based on a low dimensional setting and few guidance
exist in the context of high dimensional data, especially considering
the choice of hyperparameters. Fourth, the use of reservoir computing in
an epidemic setting with highly non-stationary time serie is scarce
whereas it is of tremendous interest for epidemiologists. Fifth, there
is currently no implementation available on R making the use of this
method challenging for user not familiar with Python.

@@

In this paper we propose to explore those challenges with extensive
guidance in order to help new user to fully benefit from RC. First a
general introduction to reservoir computing is provided with a tutorial
for its use using reservoirnet, a R package based on reservoirPy Python
module. Second, we explore the different challenges of the use of RC in
a non-stationary high dimensional setting based on the use case of
covid-19 hospitalisations forecast with extensive guidance on the
modelling strategy, the choice of hyperparameters using a genetic
algorithm and the implementation.

@@

RC is a machine learning algorithm composed of an input layer denoted
$u(t)$ which is randomy transformed by the reservoir into an activation
state denoted $x(t)$ which is fed to a ridge penalized linear regression
trained to foreast the outcome denoted as $y(t)$ as depicted at figure
\@ref(fig:rcpresentation).

@@

The input layer $u(t)$ is an $M$-dimension vector which corresponds to
the values of the input time series at time $t$ where $t = 1, …, T$. The
reservoir layer $x(t)$ is an $N_{res}$-dimensional vector where
$N_{res}$ is the number of nodes in the reservoir. The value $x(t)$ is
defined as follow:

$$x( t+1 ) = ( 1 - \alpha )  x ( t) + \alpha \: tanh( W x(t) + W_{in} u(t+1) )", "ext{ where } \alpha \in [0, 1 ]$$

The leaking rate alpha define the rate of update of the nodes. The
closer $\alpha$ is to $1$, the higher the reservoir is sensitive to new
inputs (i.e $u(t)$). Therefore, the reservoir state at time $t+1$
denoted $x(t+1)$ depends on the reservoir state at the previous time
(i.e $x(t)$) and the new inputs (i.e $u(t+1)$). Both $W_{in}$ and $W$
are random matrices of size $Nres", "imes M$ and $Nres", "imes Nres$
respectively.

$W_{in}$ is a dense matrix generated using a bernouilli distribution
where each value can be either $-I_{scale}(m)$ or $I_{scale}(m)$ with an
equal probability where $m = 1, …, M$ corresponds to a given feature in
the input layer. The input scaling, denoted $I_{scale}$, is an
hyperparameter coefficient which can be common to all features from the
input layer or specific to each feature $m$. In that case, the more
important the feature is, the greater should be its input scaling. $W$
is a sparse matrix where values are generated base on a gaussian
distribution $\mathcal{N}(0,1)$. Then the matrix $W$ is rescaled
according to the spectral radius, an hyperparameter defining the highest
eigen value of $W$.

@@

The parameter $\lambda$ is the ridge penalisation which aims to prevent
overfitting.

Overall, there are four main hyperparameters to be chosen by the user:
the leaking rate wich defines the memory of the RC, the input scaling
wich define the relative importance of the features, the spectral radius
wich define the connections of the neurons inside the reservoir and the
ridge penalization wich defines the degree of overfitting. The choice of
hyperparameter is difficult and the use of a wrapper approach where the
performance of the RC with different combinations of hyperparameters is
evaluated on a train set, the best combination is chosen for the use on
the test set.

RC in high dimension

@@

In this first use case, we will introduce the fundamental usage of the `reservoirnet` package. This demonstration will be conducted using the COVID-19 dataset that is included within the package. These data encompass hospitalization figures, positive RT-PCR results, and overall RT-PCR data sourced from Santé Publique France, which are publicly available on data.gouv.fr (for further details, refer to the `help(dfCovid)` function). Our primary objective is to predict the number of hospitalized patients 14 days into the future. To accomplish this, we will initially train our model on data preceding the date of January 1, 2022, and subsequently apply it to forecast values using the subsequent dataset.

We can proceed by loading useful packages (i.e ggplot2 [@wickham_ggplot2_2016] and dplyr [@wickham_dplyr_2023]), data and define the task:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(reservoirnet)

@@

Due to the substantial fluctuations observed in both RT-PCR metrics, our initial step involves applying a moving average computation over the most recent 7-day periods for these features. Additionally, we augment the dataset by introducing an `outcome` column and an `outcomeDate` column, which will serve as valuable inputs for model training. Moreover, we calculate the `outcome_deriv` as the difference between the outcome and the number of hospitalized patients (`hosp`), representing the variation in hospitalization in relation to the current count of hospitalized individuals. The resulting smoothed data is visualized in Figure @ref(fig:covidintro).

@@

We can now plot the data :

```{r covidintro, fig.cap="Hospitalisations, IPTCC and positive PCR of Bordeaux University Hospital.", echo = FALSE}
dfOutcome %>%
  tidyr::pivot_longer(cols = c("hosp", "Positive", "Tested")) %>%
  ggplot2::ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  theme_bw() +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  labs(color = "") +
  theme(legend.position = "bottom")
```

@@

-   Number of nodes (`units`) : it corresponds to the number of nodes inside the reservoir. Usually, the more the better but more nodes increase the computation time.
-   Leaking rate (`lr`) : the leaking rate corresponds to the balance between the new inputs and the previous state. A leaking rate of 1 only consider informations from the new inputs.
-   Spectral radius (`sr`): the spectral radius is the maximum absolute eigenvalue of the reservoir connectivity matrix. A small spectral radius induces stable dynamics inside the reservoir, a high spectral radius induces chaotic regimen inside the reservoir.
-   Input scaling (`input_scaling`): the input scaling is a gain applied to the input features of the reservoir.
-   Seed (`seed`): because the reservoir connections are set at random, setting the seed is a good approach to ensure   reproducibility. Another approach would be to use several different reservoir but it increases computation time.

For this part of the tutorial, we will set the hyperparameter at a given value. Hyperparameter opitmisation will be detailed at section [Study case](#studycase).

@@

Now we can visualise node activation using the `plot` function presented at figure \@ref(fig:nodeactivationbad).

```{r nodeactivationbad, fig.cap="Node acivation over time."}
plot(reservoir_state)
```

Numerous nodes within the system exhibit a consistent equilibrium state. The challenge arises when the ridge output layer attempts to acquire knowledge from these nodes, as they do not convey meaningful information. This issue can be attributed to the disparate scales of the features. To address this concern, a practical approach involves normalizing the features by dividing each of them by their respective maximum values, thereby rescaling them within the range of `-1` to `1`.

@@

Initially, we establish the output layer, incorporating a ridge penalty set at `1e3`. It's important to note that this hyperparameter can be subject to optimization, a topic that will be explored in the forthcoming [Study Case](#studycase) section. This parameter plays a pivotal role in fine-tuning the model's conformity to the dataset. When set excessively high, the risk of underfitting arises, whereas setting it too low can lead to overfitting. Then we connect the output layer to the reservoir making the model ready to be trained.

@@

We now standardise with the same formula as seen before. We learn the standardisation on the training set and apply it on the test set. Then we convert the dataframe to matrix.

@@

```{r fig.cap="Forecast"}
dfOutcome %>%
  mutate(pred = vec_pred) %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = pred,
                          color = "forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_bw() +
  labs(color = "", x = "Date", y = "Hospitalisations")
```

We observe that the model forecast is not fully accurate, both on the test set and the train set. In that case, one option could be to reduce ridge penalisation to fit more closely the data, the optimisation of ridge hyperparameter will be discussed at section [Study case](#studycase). Another possibility is to ease the learning of the algorithm by forecasting the variation of the hospitalisation instead of the number of hospitalised patients. For that step, we will learn on the `outcome_deriv` contained in `yTrain_variation` data which is defined outcome as `outcome_deriv = outcome - hosp`.

```{r fig.cap="Covid-19 hospitalisations forecast. The model is either trained to forecast the number of hospitalisations (denoted Raw) or the variation of the hospitalisations compared to current level of hospitalisation (denoted Variation)"}
## Fit reservoir on outcome variation instead of raw outcome
fit2 <- reservoirnet::reservoirR_fit(node = model,
                                     X = lsdf$xTrain,
                                     Y = lsdf$yTrain_variation,
                                     warmup = 30,
                                     reset = TRUE)
## Get the forecast on the test set
vec_pred2_variation <- reservoirnet::predict_seq(node = fit2$fit,
                                                 X = lsdf$xTest,
                                                 reset = TRUE)
## Transform the outome variation forecast into hospitalisation forecast
vec_pred2 <- vec_pred2_variation + xTest$hosp
## Plot the results
dfOutcome %>%
  mutate(Raw = vec_pred,
         Variation = vec_pred2) %>%
  tidyr::pivot_longer(cols = c(Raw, Variation),
                      names_to = "Outcome_type",
                      values_to = "Forecast") %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = Forecast,
                          color = "Forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  facet_wrap(Outcome_type ~ .,
             labeller = label_bquote(cols = "Outcome" : .(Outcome_type))) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_minimal() +
  labs(color = "", x = "Date", y = "Hospitalisations")
```

We observe an improvement of the forecast compared to the forecast of hospitalisation previously discussed. From there, many improvement can be implemented including : frequent update of the model, leverage additional features and hyperparameter optimisation. This notions will be discussed in the [Study case section](#studycase).

@@

This example is largely inspired from the [classification tutorial of reservoirpy](https://github.com/reservoirpy/reservoirpy/blob/master/tutorials/5-Classification-with-RC.ipynb). To illustrate the classification task, we will use the Japanese vowel dataset (@kudo_multidimensional_1999). The data can be loaded from reservoirnet as follow :

@@

The dataset comprises 640 vocalizations of the Japanese vowel \ae, contributed by nine distinct speakers. Each vocalization represents a time series spanning between 7 and 29 timesteps, encoded as a 12-dimensional vector denoting the Linear Prediction Coefficients (LPC). A visual representation of six distinct utterances from the test set, originating from three different speakers, is depicted in Figure \@ref(fig:vowelpresentation).

@@

Figure \@ref(fig:seqtovec) shows the prediction for the 6 uterrances depicted at figure \@ref(fig:vowelpresentation) where the model correctly identifies the speaker.

```{r seqtovec, fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The speaker to predict is depicted in red. For each of the 6 uterrance, the model correctly identifies the speaker."}
dfplotseqtovec <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tidyr::pivot_longer(cols = everything(),
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", "")) %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i,
                            target = speaker == pred_speaker) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtovec,
       mapping = aes(x = pred_speaker,
                     y = prediction,
                     fill = target)) +
  geom_bar(stat = "identity") +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  scale_fill_manual(values = c("darkgrey", "darkred")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y = 'Score',
       x = "Speaker")
```

@@

# Study case: Covid-19 hospitalisations forecast {#studycase}

@@

Since late 2020, millions of cases of SARS-CoV-2 infection have been documented across the globe [@carrat_evidence_2021, @world_health_organisation_who_2020, @noauthor_estimating_2022]. This ongoing pandemic has exerted significant strain on healthcare systems, resulting in a surge in hospitalizations. This surge, in turn, necessitated modifications to the healthcare infrastructure and gave rise to unprecedented population-wide lockdown measures aimed at preventing the saturation of healthcare facilities [@kim_health_2020, @simoes_organisation_2021, @hubner_surgery_2020]. The capacity to predict the trajectory of the epidemic on a regional scale is of paramount importance for effective healthcare system management.

Numerous COVID-19 forecasting algorithms have been proposed using different methods (e.g ensemble, deep learning, compartmental), yet none has proven entirely satisfactory [@rahimi_review_2021, @cramer_evaluation_2022]. In France, short-term forecasts  with different methods have been evaluated with similar results [@pottier_forecast_2021, @paireau_ensemble_2022, @carvalho_analysis_2021, @mohimont_convolutional_2021]. In this context a machine learning algorithm based on linear regression with elastic-net penalisation, leveraging both EHRs and public data, was implemented at Bordeaux University Hospital [@ferte_benefit_2022]. This model, which aimed at forecasting the number of hospitalised patients at 14 days, showed good performance but struggled to accurately anticipate dynamic shifts of the epidemic.

RC has been used in the context of covid-19 epidemic forecast [@ghosh_reservoir_2021, @kmet_bezier_2019, @liu_nanophotonic_2023, @ray_optimized_2021, @zhang_sentiment_2023]. Among them, @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021 used it to forecast epidemic (@zhang_sentiment_2023 performed sentiment analysis and @kmet_bezier_2019 used it to solve optimal control related to vaccine). The evaluation of RC for epidemic forecast showed promising results in all approaches, being competitive with Long-Short Term Memory (LSTM) and Feed-Forward Neural Network (FFNN) in @ray_optimized_2021. However, the test period was short for @ghosh_reservoir_2021 (21 and 14 days) and @ray_optimized_2021 (86 days) making it difficult to evaluate the behavior of the methods during epidemic dynamic shift. This was not the case for @liu_nanophotonic_2023 (6 months) but they implemented daily ahead forecast which would be difficult to use to manage a hospital. Finally, all three implementations used only one time series as input whereas it has been shown that using different data sources could improve forecast [@ferte_benefit_2022]. Therefore, it is still difficult to assess the usefulness of RC over a large period and using many time series as inputs.

@@

`TODO :` add paragraph about hp optimisation for RC (focus on GA)

RC can be approached as a complexification of penalized linear regression where inputs are processed by a reservoir which allows memory and non-linear combinations. As penalised linear regression was the best approach for covid-19 forecast in @ferte_benefit_2022 and that RC showed promising results for epidemic forecast in @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021, we proposed to use RC to forecast hospitalisations at 14 days at the University Hospital of Bordeaux. The main objective of this study is to evaluate the performance of RC for this task. Secondary objective will be (i) to assess the performance of RC depending on the input dimension, (ii) to assess the relevance of Genetic Algorithm (GA) for hyperparameter optimisation and feature selection in this context.

@@

The task was to forecast 14 days ahead the number of hospitalised patients. As seen at section [Regression](#basicregression), we will train the model to predict the variation of hospitalisation defined as $outcome_{t+14} = hosp_{t+14} - hosp_{t}$. Metrics computation and visualisations will be performed on the predicted number of hospitalizations denoted as $\hat{y_{t+14}} = \widehat{outcome_{t+14}} + hospitalizations_{t}$ and the observed number of hospitalisations at time $t+14$ denoted as $y_{t+14} = hospitalizations_{t+14}$

The dataset was separated into two periods. First period from May 16, 2020 to March 1, 2021 served to identify relevant hyperparameters. Remaining data was used to evaluate the model performance. As the time serie is not stationary and most appropriate hyperparameter value might change over time, we also investigated the effect of hyperparameter update every month.

>>>>

**TODO : cite : Jaeger, Herbert. “The “echo state” approach to analysing and training recurrent neural networks-with an erratum note.” Bonn, Germany: German National Research Center for Information Technology GMD Technical Report 148.34 (2001): 13.**
**Maass W, Natschläger T, Markram H. Real-time computing without stable states: a new framework for neural computation based on perturbations. Neural Comput. 2002;14(11):2531‐2560.**

@@

RC exhibits several advantages over its counterparts. It excels in tasks
that require the processing of temporal data, such as time series
forecasting and speech recognition, thanks to its inherent memory
capabilities. RC offers computational efficiency and reduced susceptibility
to overfitting compared to deep learning. Moreover, RC is able to capture intricate temporal
patterns compared to statistical methods such as penalized regression.

**TODO : find ref to support it.**

@@

**TODO: add ref from use case intro on use in epidemiology + find review**
**Table 1 contains a list of use case of RC : https://www.sciencedirect.com/science/article/pii/S0893608019300784**

@@

RC use is still challenging for several reasons. First, its stochastic
nature due to the reservoir random part makes the results of a given RC
unpredictable even if the hyperparameters are well chosen. Second,
even though, RC relies on several hyperparameter to determine the
connections between neurons. There exist few guidance about the way to
set the hyperparameters and most often the user must rely on a wrapper
approach were combinations of hyperparameters performance are evaluated
on a train set before choosing the ideal combination one. Third, most
published paper are based on a low dimensional setting and few guidance
exist in the context of high dimensional data, especially considering
the choice of hyperparameters. Fourth, the use of reservoir computing in
an epidemic setting with highly non-stationary time series is scarce
whereas it is of tremendous interest for epidemiologists. Fifth, there
is currently no implementation available on R making the use of this
method challenging for user not familiar with Python.

**TODO : extend this part and link to previous work**
**Incorporate the following :**
RC in high dimension

@@

In this paper we propose to explore those challenges with extensive
guidance in order to help new user to fully benefit from RC. First a
general introduction to reservoir computing is provided with a tutorial
for its use using `reservoirnet`, a R package based on `reservoirPy` Python
module. Second, we explore the different challenges of the use of RC in
a non-stationary high dimensional setting based on the use case of
covid-19 hospitalizations forecast with extensive guidance on the
modelling strategy, the choice of hyperparameters using a genetic
algorithm and the implementation.

@@

RC is a machine learning algorithm composed of an input layer denoted
$u(t)$ which is randomly transformed by the reservoir into an activation
state denoted $x(t)$ which is fed to a ridge penalized linear regression
trained to forecast the outcome denoted as $y(t)$ as depicted at figure
\@ref(fig:rcpresentation).

@@

The input layer $u(t)$ is an $M$-dimension vector, where $M$ is the number of input time series, which corresponds to the values of the input time series at time $t$ where $t = 1, …, T$. The reservoir layer $x(t)$ is an $N_{res}$-dimensional vector where $N_{res}$ is the number of nodes in the reservoir. The value $x(t)$ is defined as follow:

$$x( t+1 ) = ( 1 - \alpha )  x ( t) + \alpha \: tanh( W x(t) + W_{in} u(t+1) ) \text{ , where } \alpha \in [0, 1 ]$$

The leaking rate alpha define the rate of update of the nodes. The
closer $\alpha$ is to $1$, the higher the reservoir is sensitive to new
inputs (i.e $u(t)$). Therefore, the reservoir state at time $t+1$
denoted $x(t+1)$ depends on the reservoir state at the previous time
(i.e $x(t)$) and the new inputs (i.e $u(t+1)$). Both $W_{in}$ and $W$
are random matrices of size $Nres \times M$ and $Nres \times Nres$
respectively.

$W_{in}$ is a dense matrix generated using a Bernoulli distribution
where each value can be either $-I_{scale}(m)$ or $I_{scale}(m)$ with an
equal probability where $m = 1, …, M$ corresponds to a given feature in
the input layer. The input scaling, denoted $I_{scale}$, is an
hyperparameter coefficient which can be common to all features from the
input layer or specific to each feature $m$. In that case, the more
important the feature is, the greater should be its input scaling. $W$
is a sparse matrix where values are generated from a Gaussian
distribution $\mathcal{N}(0,1)$. Then the matrix $W$ is scaled
according to the spectral radius, an hyperparameter defining the highest
eigen value of $W$.

@@

The parameter $\lambda$ is the ridge penalization which aims to prevent
overfitting.

Overall, there are four main hyperparameters to be chosen by the user: the leaking rate which defines the memory of the RC, the input scaling which define the relative importance of the features, the spectral radius which define the connections of the neurons inside the reservoir which in turn define the degree of non-linear combination of features and the ridge penalization which controls the degree of overfitting. The choice of hyperparameter is difficult and the use of a wrapper approach where the performance of the RC with different combinations of hyperparameters is evaluated on a train set, the best combination is chosen for the use on the test set.

@@

In this first use case, we will introduce the fundamental usage of the `reservoirnet` package. This demonstration will be conducted using the COVID-19 dataset that is included within the package. These data encompass hospitalization, positive RT-PCR results, and overall RT-PCR data sourced from Santé Publique France, which are publicly available on data.gouv.fr (for further details, refer to `help(dfCovid)`). Our primary objective is to predict the number of hospitalized patients 14 days into the future. To accomplish this, we will initially train our model on data preceding the date of January 1, 2022, and subsequently apply it to forecast values using the subsequent dataset.

We can proceed by loading useful packages (i.e ggplot2 [@wickham_ggplot2_2016] and dplyr [@wickham_dplyr_2023], **TODO : add cowplot**), data and define the task:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(reservoirnet)
library(cowplot)

@@

Due to the substantial fluctuations observed in both RT-PCR metrics, our initial step involves applying a moving average computation over the most recent 7-day periods for these features. Additionally, we augment the dataset by introducing an `outcome` column and an `outcomeDate` column, which will serve as valuable inputs for model training. Moreover, we calculate the `outcome_deriv` as the difference between the outcome and the number of hospitalized patients (`hosp`), representing the variation in hospitalization in relation to the current count of hospitalized individuals. The resulting smoothed data is visualized in Figure \@ref(fig:covidintro).

@@

```{r covidintro, fig.cap="Hospitalizations, IPTCC and positive PCR of Bordeaux University Hospital.", echo = FALSE}
dfOutcome %>%
  tidyr::pivot_longer(cols = c("hosp", "Positive", "Tested")) %>%
  ggplot2::ggplot(mapping = aes(x = date, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  theme_bw() +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  labs(color = "") +
  theme(legend.position = "bottom")
```

@@

-   Number of nodes (`units`) : it corresponds to the number of nodes inside the reservoir. Usually, the more the better but more nodes increases the computation time.
-   Leaking rate (`lr`) : the leaking rate corresponds to the balance between the new inputs and the previous state. A leaking rate of 1 only consider information from new inputs.
-   Spectral radius (`sr`): the spectral radius is the maximum absolute eigenvalue of the reservoir connectivity matrix. A small spectral radius induces stable dynamics inside the reservoir, a high spectral radius induces chaotic regimen inside the reservoir.
-   Input scaling (`input_scaling`): the input scaling is a gain applied to the input features of the reservoir.
-   Seed (`seed`): because the reservoir connections are set at random, setting the seed is a good approach to ensure   reproducibility. Another approach would be to use several different reservoir but it increases computation time.

For this part of the tutorial, we will set the hyperparameter at a given value. Hyperparameter optimization will be detailed at section [Study case](#studycase).

@@

Now we can visualize node activation using the `plot` function presented at figure \@ref(fig:nodeactivationbad).

```{r nodeactivationbad, fig.cap="Node activation over time."}
plot(reservoir_state)
```

Numerous nodes within the system exhibit a consistent equilibrium state. The challenge arises when the ridge output layer attempts to acquire knowledge from these nodes, as they do not convey meaningful information. This issue can be attributed to the disparate scales of the features. To address this concern, a practical approach involves normalizing the features by dividing each of them by their respective maximum values, thereby scaling them within the range of `-1` to `1`.

@@

Initially, we establish the output layer, incorporating a ridge penalty set at `1e3`. It's important to note that this hyperparameter can be subject to optimization, a topic that will be explored in the forthcoming [Study Case](#studycase) section. This parameter plays a pivotal role in fine-tuning the model's conformity to the data. When set excessively high, the risk of underfitting arises, whereas setting it too low can lead to overfitting. We connect the output layer to the reservoir making the model ready to be trained.

@@

We standardize with the same formula as seen before. We learn the standardization on the training set and apply it on the test set. Then we convert the dataframe to matrix.

@@

```{r fig.cap="Forecast"}
dfOutcome %>%
  mutate(pred = vec_pred) %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = pred,
                          color = "forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_bw() +
  labs(color = "", x = "Date", y = "Hospitalizations")
```

We observe that the model forecast is not fully accurate, both on the test set and the train set. In that case, one option could be to reduce ridge penalization to fit more closely the data, the optimization of ridge hyperparameter will be discussed at section [Study case](#studycase). Another possibility is to ease the learning of the algorithm by forecasting the variation of the hospitalization instead of the number of hospitalized patients. For that step, we will learn on the `outcome_deriv` contained in `yTrain_variation` data which is defined outcome as `outcome_deriv = outcome - hosp`.

```{r fig.cap="Covid-19 hospitalizations forecast. The model is either trained to forecast the number of hospitalizations (denoted Raw) or the variation of the hospitalizations compared to current level of hospitalisation (denoted Variation)"}
## Fit reservoir on outcome variation instead of raw outcome
fit2 <- reservoirnet::reservoirR_fit(node = model,
                                     X = lsdf$xTrain,
                                     Y = lsdf$yTrain_variation,
                                     warmup = 30,
                                     reset = TRUE)
## Get the forecast on the test set
vec_pred2_variation <- reservoirnet::predict_seq(node = fit2$fit,
                                                 X = lsdf$xTest,
                                                 reset = TRUE)
## Transform the outome variation forecast into hospitalisation forecast
vec_pred2 <- vec_pred2_variation + xTest$hosp
## Plot the results
dfOutcome %>%
  mutate(Raw = vec_pred,
         Variation = vec_pred2) %>%
  tidyr::pivot_longer(cols = c(Raw, Variation),
                      names_to = "Outcome_type",
                      values_to = "Forecast") %>%
  na.omit() %>%
  ggplot(mapping = aes(x = outcomeDate)) +
  geom_line(mapping = aes(y = outcome,
                          color = "observed")) +
  geom_line(mapping = aes(y = Forecast,
                          color = "Forecast")) +
  geom_vline(mapping = aes(color = "train-test sets",
                           xintercept = traintest_date)) +
  facet_wrap(Outcome_type ~ .,
             labeller = label_bquote(cols = "Outcome" : .(Outcome_type))) +
  scale_color_manual(values = c("#3772ff", "#080708", "#df2935")) +
  theme_minimal() +
  labs(color = "", x = "Date", y = "Hospitalizations")
```

We observe an improvement of the forecast compared to the forecast of hospitalization previously discussed. From there, many improvement can be implemented including : frequent update of the model, leverage additional features and hyperparameter optimization. This notions will be discussed in the [Study case section](#studycase).

@@

This example is largely inspired from the [classification tutorial of reservoirpy](https://github.com/reservoirpy/reservoirpy/blob/master/tutorials/5-Classification-with-RC.ipynb). To illustrate the classification task, we will use the Japanese vowel dataset (@kudo_multidimensional_1999). The data can be loaded from `reservoirnet` as follow :

@@

The dataset comprises 640 vocalizations of the Japanese vowel \ae, contributed by nine distinct speakers. Each vocalization represents a time series spanning between 7 and 29 time steps, encoded as a 12-dimensional vector denoting the Linear Prediction Coefficients (LPC). A visual representation of six distinct utterances from the test set, originating from three different speakers, is depicted in Figure \@ref(fig:vowelpresentation).

@@

Figure \@ref(fig:seqtovec) shows the prediction for the 6 utterances depicted at figure \@ref(fig:vowelpresentation) where the model correctly identifies the speaker.

```{r seqtovec, fig.cap="Prediction in a sequence-to-sequence approach 6 samples with 3 speakers and 2 utterance each. The speaker to predict is depicted in blue. For each of the 6 utterance, the model correctly identifies the speaker."}
dfplotseqtovec <- lapply(vec_sample,
                 FUN = function(i){
                   speaker <- which(Y_test[[i]][1,] == 1)
                   Y_pred[[i]] %>%
                     as.data.frame() %>%
                     tidyr::pivot_longer(cols = everything(),
                                         names_to = "pred_speaker",
                                         values_to = "prediction") %>%
                     mutate(pred_speaker = gsub(x = pred_speaker,
                                                pattern = "V", "")) %>%
                     mutate(speaker = speaker, .before = 1,
                            uterrance = i,
                            target = speaker == pred_speaker) %>%
                     return()
                 }) %>%
  bind_rows()

ggplot(dfplotseqtovec,
       mapping = aes(x = pred_speaker,
                     y = prediction,
                     fill = target)) +
  geom_bar(stat = "identity") +
  facet_wrap(uterrance ~ speaker,
             labeller = label_bquote(cols = "speaker" : .(speaker)),
             ncol = 2) +
  scale_fill_manual(values = c("#BDBDBD", "#A3CEF1")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y = 'Score',
       x = "Speaker")
```

@@

# Study case: Covid-19 hospitalizations forecast {#studycase}

@@

Since late 2020, millions of cases of SARS-CoV-2 infection have been documented across the globe [@carrat_evidence_2021, @world_health_organisation_who_2020, @noauthor_estimating_2022]. This ongoing pandemic has exerted significant strain on healthcare systems, resulting in a surge in hospitalizations. This surge, in turn, necessitated modifications to the healthcare infrastructure and gave rise to population-wide lockdown measures aimed at preventing the saturation of healthcare facilities [@kim_health_2020, @simoes_organisation_2021, @hubner_surgery_2020]. The capacity to predict the trajectory of the epidemic on a regional scale is of paramount importance for effective healthcare system management.

Numerous COVID-19 forecasting algorithms have been proposed using different methods (e.g ensemble, deep learning, compartmental), yet none has proven entirely satisfactory [@rahimi_review_2021, @cramer_evaluation_2022]. In France, short-term forecasts  with different methods have been evaluated with similar results [@pottier_forecast_2021, @paireau_ensemble_2022, @carvalho_analysis_2021, @mohimont_convolutional_2021]. In this context a machine learning algorithm based on linear regression with elastic-net penalization, leveraging both Electronic Health Records (EHRs) and public data, was implemented at Bordeaux University Hospital [@ferte_benefit_2022]. This model, which aimed at forecasting the number of hospitalized patients at 14 days, showed good performance but struggled to accurately anticipate dynamic shifts of the epidemic.

RC has been used in the context of covid-19 epidemic forecast [@ghosh_reservoir_2021, @kmet_bezier_2019, @liu_nanophotonic_2023, @ray_optimized_2021, @zhang_sentiment_2023]. Among them, @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021 used it to forecast epidemic, @zhang_sentiment_2023 performed sentiment analysis and @kmet_bezier_2019 used it to solve optimal control related to vaccine. The evaluation of RC for epidemic forecast showed promising results in all approaches, being competitive with Long-Short Term Memory (LSTM) and Feed-Forward Neural Network (FFNN) in @ray_optimized_2021. However, the test period was short for @ghosh_reservoir_2021 (21 and 14 days) and @ray_optimized_2021 (86 days) making it difficult to evaluate the behavior of the methods during epidemic dynamic shift. This was not the case for @liu_nanophotonic_2023 (6 months) but they implemented daily ahead forecast which would be difficult to use to manage a hospital. Finally, all three implementations used only one time series as input whereas it has been shown that using different data sources could improve forecast [@ferte_benefit_2022]. Therefore, it is still difficult to assess the usefulness of RC over a large period and using many time series as inputs.

`TODO :` check for new published forecast models

@@

`TODO :` add paragraph about hp optimisation for RC (focus on GA). Discussed in https://ieeexplore.ieee.org/abstract/document/8481347

RC can be approached as an extension of penalized linear regression where inputs are processed by a reservoir which allows memory and non-linear combinations. As penalized linear regression was the best approach for covid-19 forecast in @ferte_benefit_2022 and that RC showed promising results for epidemic forecast in @ghosh_reservoir_2021, @liu_nanophotonic_2023 and @ray_optimized_2021, we proposed to use RC to forecast hospitalizations at 14 days at the University Hospital of Bordeaux. The main objective of this study is to evaluate the performance of RC for this task. Secondary objective will be (i) to assess the performance of RC depending on the input dimension, (ii) to assess the relevance of Genetic Algorithm (GA) for hyperparameter optimization and feature selection in this context.

@@

The task was to forecast 14 days ahead the number of hospitalized patients. As seen at section [Regression](#basicregression), we will train the model to predict the variation of hospitalization defined as $outcome_{t+14} = hosp_{t+14} - hosp_{t}$. Metrics computation and visualizations will be performed on the predicted number of hospitalizations denoted as $\hat{y_{t+14}} = \widehat{outcome_{t+14}} + hospitalizations_{t}$ and the observed number of hospitalizations at time $t+14$ denoted as $y_{t+14} = hospitalizations_{t+14}$

The dataset was separated into two periods. First period from May 16, 2020 to March 1, 2021 served to identify relevant hyperparameters. Remaining data was used to evaluate the model performance. As the time series is not stationary and most appropriate hyperparameter value might change over time, we also investigated the effect of hyperparameter update every month.
